Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2022-05-06 article/draft.tex
typos

Hence, in the present article, we focus on mainstream platforms' policies and interventions regarding content with low credibility or false information, commonly referred to as {\it Fake News} (see Lazer et al. (2018)~\cite{lazer}). The misinformation phenomenon is still ill-defined by the academic community, as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience.  The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society when it goes viral. This gave rise to a set of heterogenous fact-checking policies across mainstream platforms. For example, Facebook has a substantial partnership program with Fact-checking partners certified by the non-partisan International Fact-Checking Network. The company uses a number of signals and machine learning models to predict misinformation and surface it to fact-checkers.\footnote{See the \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{section Frequently asked questions: `How does Facebook use technology to detect potential misinformation?''}} Twitter seems to have a different approach where they focus on providing context rather than fact-checking\footnote{To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} and the platform is testing a new system based on the wisdom of the crowds to tackle misinformation (see \href{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}{Twitter Birdwatch}). As for YouTube, this platform utilizes the \href{https://schema.org/ClaimReview}{schema.org ClaimReview} markup, where fact-checking articles created by eligible publishers can appear on information panels (see Appendix \ref{links} for references).

@@

During the COVID-19 global health pandemic, platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content (see Appendix \ref{links} for references). Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of mainstream platforms to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. The 2019 report of the Facebook Data Transparency Advisory Group (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report}''. In particular, policies regarding misinformation are generally not part of the set of platform rules or community guidelines (as of July 2021).  %Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

@@

Misinformation related interventions by mainstream platforms are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify 
mainstream platforms' current actions regarding content with low credibility or false information
with data mining. We do so by providing a series of examples for different interventions and platforms. We chose to focus on three platforms: Facebook, Twitter and YouTube. Both Facebook and YouTube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. To collect data from these three platforms, we either used the APIs (Application Programming Interfaces), or web scraping, i.e. retro-engineering the HMTL code of a web page to extract meaningful data, see Table \ref{tab1} for a summary. Minet \cite{minet}, a webmining tool developed by the SciencesPo médialab, was often used, and we scripted our own data mining code when it was necessary.

@@

Mainstream platforms such as Facebook, Twitter and Youtube are increasingly working towards higher data transparency (see links that redirect to the respective transparency centers in Appendix \ref{links}). 
Yet the present research sheds light on the restrictions to access data and its limits, which impede the academic community from studying successfully online misinformation and assessing the impact of platforms' interventions regarding this phenomenon. 
Hence, few comments are in order regarding the limitations to access data, the lack of transparency vis-\`{a}-vis misinformation related interventions and the lack of interventions designed specifically for misinformation.

@@

Although it is currently impossible to verify permanent deletions using available social media data, their effects can be observed using an adapted methodology.
Following the Capitol riots in January $2021$, Twitter has announced the suspension of more than $70$ $000$ accounts related to QAnon.\footnote{\href{https://blog.twitter.com/en\_us/topics/company/2021/protecting--the-conversation-following-the-riots-in-washington--}{https://blog.twitter.com/en\_us/topics/company/2021/protecting--the-conversation-following-the-riots-in-washington--}.} 
A journalist has recently investigated the effects of this measure by counting the number of times hashtags related to the QAnon movement have been used over time, such as \#QAnon or \#WWG1WGA.
A drastic reduction in hashtag use was found when comparing the first quarter of $2021$ with the first quarter of $2020$.\footnote{See \href{https://twitter.com/Shayan86/status/1394742784683298818}{https://twitter.com/Shayan86/status/1394742784683298818}.}
Thus, counting specific hashtag use over time allows to indirectly observe permanent suspensions on Twitter and there is a need for similar methodologies to be experimented on data from other platforms.
However, a point of concern is that data can also disappear because of other factors independent of platforms' content moderation.
To provide an example, McMinn et al. (2012)\cite{mcminn2013building} collected $120$ million Tweets related to general topics. 
The same published set of Tweet IDs was then used by Mazoyer et al. (2020)~\cite{mazoyer2020french}, to run the same collection in $2018$. Mazoyer et al. (2020)~\cite{mazoyer2020french} were only able to retrieve $67$ millions of Tweets (55\%). The authors argue that this drop is probably due to Tweets having been removed by users and Twitter accounts being shut by the users themselves.
As data can disappear from platforms over time, it is hard to distinguish whether platforms or users are the ones responsible for data removal. 
Hence in the special case of Twitter accounts related to QAnon, it is hard to identify how much of the drop in QAnon related hashtag use is due to the Twitter intervention or to accounts migrating to other platforms out of fear of having their accounts suspended by the platform.

@@

A last concern regards the growing restrictions to access APIs (Perriam et al.(2019)~\cite{api}).
In April $2018$, likely as a response to the Cambridge Analytica scandal, Facebook severely limited access to their official APIs\footnote{See \href{https://about.fb.com/news/2018/04/restricting-data-access/}{https://about.fb.com/news/2018/04/restricting-data-access/} and \href{http://thepoliticsofsystems.net/2018/08/facebooks-app-review-and-how-independent-research-just-got-a-lot-harder/}{http://thepoliticsofsystems.net/2018/08/} \href{http://thepoliticsofsystems.net/2018/08/facebooks-app-review-and-how-independent-research-just-got-a-lot-harder/}{/facebooks-app-review-and-how-independent-research-just-got-a-lot-harder/}.} 
and CrowdTangle is today a reliable source for researchers to access Facebook data. 
However, it should be noted that CrowdTangle has been recently used by journalists to show that Facebook is disproportionally favoring highly partisan information\footnote{See the news article \href{https://www.economist.com/graphic-detail/2020/09/10/facebook-offers-a-distorted-view-of-american-news}{https://www.economist.com/graphic-detail/2020/09/10/facebook-offers-a-distorted-view-of-american-news} and the Tweets from \href{https://twitter.com/facebookstop10}{https://twitter.com/facebookstop10}. 
The last link is a Twitter account showing each day the ten most shared link on Facebook, most of them coming from extreme right and misinformation sources. 
Facebook has reacted to this here: \href{https://about.fb.com/news/2020/11/what-do-people-actually-see-on-facebook-in-the-us/}{https://about.fb.com/news/2020/11/what-do-people-actually-see-on-facebook-in-the-us/}.} 
and a general suspicion is growing that its access might be shut down in a near future to journalists and researchers.\footnote{See the news articles \href{https://www.nytimes.com/2021/07/14/technology/facebook-data.html}{https://www.nytimes.com/2021/07/14/technology/facebook-data.html} and \href{https://www.bloomberg.com/news/articles/2021-08-03/facebook-disables-accounts-tied-to-nyu-research-project}{https://www.bloomberg.com/news/articles/2021-08-03/facebook-disables-accounts-tied-to-nyu-research-project}.}
Regarding Twitter and YouTube, it is still possible and relatively easy to access their official APIs. Nevertheless the number of Tweets that can be retrieved has decreased between the Twitter API v1 and v2. 
In the v2, there is a monthly cap of $10,000,000$ Tweets for the academic research track\footnote{https://developer.twitter.com/en/docs/projects/overview},
while in the v1 there was no monthly limit and only a limit on the number of requests that could be made in a 15-minute window\footnote{https://developer.twitter.com/en/docs/twitter-api/v1/rate-limits}.
We thus noticed that we could get around $10$ times more Tweets per month with the v1 when compared to the v2 for one specific collection.

@@

In certain cases, the platform interventions were not motivated by content moderation to limit misinformation, but rather by other rule violations.
For example, the official reason for the suspension of Donald Trump's Twitter account was not misinformation\footnote{Although Donald Trump may have spread many misinformation during his governance, see for example: \href{http://allianceforscience.cornell.edu/blog/2020/10/what-drove-the-covid-misinformation-infodemic/}{http://allianceforscience.cornell.edu/blog/2020/10/what-drove-the-covid-misinformation-infodemic/}.}
but the violation of Twitter's Glorification of Violence Policy, with his last Tweets encouraging the Capitol's riot.\footnote{See \href{https://blog.twitter.com/en\_us/topics/company/2020/suspension}{https://blog.twitter.com/en\_us/topics/company/2020/suspension}.}
Similarly, when Facebook removed four pages related to the InfoWars website, the cited reason was repeated violations of Community Standards: 
``While much of the discussion around Infowars has been related to false news, which is a serious issue that we are working to address [...], none of the violations that spurred today’s removals were related to this''.\footnote{See the article \href{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}.}
Regarding The Beauty of Life website, its ban was officially due to coordinated inauthentic behavior,\footnote{See \href{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us/}{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us/}.} 
which includes using fake accounts that misrepresent one's identity or using methods to artificially boost the popularity of content. 
According to Facebook, coordinated inauthentic behavior is a distinct phenomenon from disinformation, as ``most of the content shared by coordinated manipulation campaigns isn’t probably false''.\footnote{See: \href{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}.} 
Nevertheless, both misinformation and coordinated inauthentic behavior were attested for the Beauty of Life, and reported to Facebook by the fact-checking organization Snopes.\footnote{See \href{ https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}, 
\href{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-} \href{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{fake-profiles/}  and 
\href{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}.} 
In the above ``known'' examples, the compagnies' communication was very clear on the reason of the interventions (although one can suspect that the official reason given can depart from the real motivation behind an intervention). 
However, very often no official communication on behalf of the platforms can be found for a specific interventions observed in collected data. 
Because misinformation is often associated with creating fake accounts to speed its spread, and incitation to violence, it is sometimes unclear to distinguish between interventions due to misinformation and interventions due to other guideline violations, and more transparent communication on platforms' actions would help monitor the interventions specifically related to misinformation.

@@

{\bf A need for more consistent and adapted policies.} Misinformation on the web is a problem that is increasingly discussed since the $2016$ American elections and in the years that followed, platforms may have decided on a case-by-case basis which actions to enforce.
Infowars and The Beauty of Life domains have both several failed fact-checks\footnote{See \href{https://mediabiasfactcheck.com/infowars-alex-jones/}{https://mediabiasfactcheck.com/infowars-alex-jones/}  and \href{https://mediabiasfactcheck.com/the-bl/}{https://mediabiasfactcheck.com/the-bl/}.} and have both violated Facebook's community guidelines (incitation to violence and coordinated unauthentic behavior). 
Yet one's distribution was reduced for only half of a year, while the other one's links cannot be shared anymore on Facebook since the end of $2019$, and it is not clear why such different enforcements were applied to these comparable websites.
It is also possible that platforms usually follow specific rules on how to fight misinformation, but that sometimes the sudden media coverage of specific events leads them to take exceptional actions, such as with the Capitol riots and Trump's suspension from Facebook, Twitter and YouTube.
We would argue that to efficiently curb the spread of misinformation, policies should be clearly stated and consistently applied, rather than implementing interventions on a case-by-case basis and in response to the latest news events.

@@

Furthermore, it is likely that most interventions used to tackle online misinformation correspond to interventions that were designed for earlier policy areas.
To illustrate, YouTube explicitly states
\footnote{See \href{https://www.youtube.com/intl/en\_us/howyoutubeworks/our-commitments/fighting-misinformation/\#policies}{What policies exist to fight misinformation on Youtube?}.}
that: ``Several policies in our Community Guidelines are directly applicable to misinformation'', such as the deceptive practices, impersonation and hate speech policies.
Hence interventions such as content deletion or account suspension, might be efficient for policy areas for which they were specifically designed, such as the policy area {\it Adult Nudity \& Sexual Activity}. Nevertheless, tackling online misinformation might need adapted interventions.
Namely, account suspensions or content deletion motivated by the spread of inaccurate information can be perceived as censorship by the targeted users. In particular, such interventions can generate platform migrations (see Rogers (2020)~\cite{rogers2020}) towards new alternative social media platforms (e.g. Telegram, Gab, Minds) with little or no content moderation.

Finally, the means to tackle online misinformation is an active area of research. Developing effective interventions against misinformation depends on understanding the interventions currently used by platforms (e.g. for labeling, see Morrow et al. (2021)~\cite{morrow}) along with the psychology of interaction with social media (Pennycook and Rand (2021)~\cite{pennycook2021psychology}). 
Adding fact-checking labels to controversial content is certainly the intervention whose effectiveness was the most studied so far. 
It was shown that credibility indicators do reduce misperceptions and sharing, although their impact varied vastly depending on the presentation of the warnings (see Yaqub (2020)~\cite{yaqub2020effects} and Sharevski (2021)~\cite{sharevski2021misinformation}).
Moreover, a fact-checking label can lead to unpredicted consequences depending on the perception of the person reading it.
On Facebook, Twitter and YouTube, fact-checking messages are currently applied only to some content, while the rest of the content on a given platform remains naturally unlabelled. 
Such sparse application of warnings can lead to an ``implied truth'' effect, where users may assume that content without warnings have actually been verified and shown to be true (Pennycook et al. (2020)~\cite{pennycook2020implied}).
Using a different approach, Mosleh et al. (2021)~\cite{mosleh2021perverse} identified Twitter users sharing false news and replied to their false Tweets with links redirecting to fact-checking websites.
They observed a decrease in the quality and accuracy of the content subsequently shared by the corrected user. 
These research papers highlight the necessity to investigate potential unintended effects of well-meaning interventions. Furthermore, misinformation related interventions can be poorly misunderstood. Based on survey data, Saltz et al. (2021)~\cite{saltzbarrari} show that $40\%$ of the survey respondents (wrongly) believed that content is mostly or all fact-checked. To date, research about misinformation related interventions by platforms is still scarce and addressing the specific needs in terms of data for auditing and monitoring tasks can greatly help in improving our understanding of these interventions and study their impact.  
The methods presented in this paper can help to monitor not only the platforms' actions against misinformation, but also their subsequent consequences in real-life settings.

>>>>

Hence, in the present article, we focus on mainstream platforms' policies and interventions regarding content with low credibility or false information, commonly referred to as {\it Fake News} (see Lazer et al. (2018)~\cite{lazer}). The misinformation phenomenon is still ill-defined by the academic community, as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience.  The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society when it goes viral. This gave rise to a set of heterogenous policies across mainstream platforms, mainly based on fact-checking but also content moderators and users' reporting. For example, Facebook has a substantial partnership program with Fact-checking partners certified by the non-partisan International Fact-Checking Network. The company uses a number of signals and machine learning models to predict misinformation and surface it to fact-checkers.\footnote{See the \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{section Frequently asked questions: `How does Facebook use technology to detect potential misinformation?''}} Twitter seems to have a different approach where they focus on providing context rather than fact-checking\footnote{To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} and the platform is testing a new system based on the wisdom of the crowds to tackle misinformation (see \href{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}{Twitter Birdwatch}). As for YouTube, this platform utilizes the \href{https://schema.org/ClaimReview}{schema.org ClaimReview} markup, where fact-checking articles created by eligible publishers can appear on information panels (see Appendix \ref{links} for references).

@@

During the COVID-19 global health pandemic, platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content (see Appendix \ref{links} for references). Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those policies are generally not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of mainstream platforms to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. The 2019 report of the Facebook Data Transparency Advisory Group (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report}''. In particular, policies regarding misinformation are generally not part of the set of platform rules or community guidelines (as of July 2021).  %Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

@@

Misinformation related interventions by mainstream platforms are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and to date, it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify 
mainstream platforms' current actions regarding content with low credibility or false information
with data mining. We do so by providing a series of examples for different interventions and platforms. We chose to focus on three platforms: Facebook, Twitter and YouTube. Both Facebook and YouTube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. To collect data from these three platforms, we either used the APIs (Application Programming Interfaces), or web scraping, i.e. retro-engineering the HMTL code of a web page to extract meaningful data, see Table \ref{tab1} for a summary. Minet \cite{minet}, a webmining tool developed by the SciencesPo médialab, was often used, and we scripted our own data mining code when it was necessary.

@@

Mainstream platforms such as Facebook, Twitter and Youtube are increasingly working towards higher data transparency. We invite the reader to see links that redirect to the respective transparency centers in Appendix \ref{links}. 
Yet the present research sheds light on the restrictions to access data and its limits, which impede the academic community from studying successfully online misinformation and assessing the impact of platforms' interventions regarding this phenomenon. 
Hence, few comments are in order regarding the limitations to access data, the lack of transparency vis-\`{a}-vis misinformation related interventions and the lack of interventions designed specifically for misinformation.

@@

Although it is currently impossible to verify permanent deletions using available social media data, their effects can be observed using an adapted methodology.
Following the Capitol riots in January $2021$, Twitter has announced the suspension of more than $70$ $000$ accounts related to QAnon.\footnote{\href{https://blog.twitter.com/en\_us/topics/company/2021/protecting--the-conversation-following-the-riots-in-washington--}{https://blog.twitter.com/en\_us/topics/company/2021/protecting--the-conversation-following-the-riots-in-washington--}.} 
A journalist has recently investigated the effects of this measure by counting the number of times hashtags related to the QAnon movement have been used over time, such as \#QAnon or \#WWG1WGA.
A drastic reduction in hashtag use was found when comparing the first quarter of $2021$ with the first quarter of $2020$.\footnote{See \href{https://twitter.com/Shayan86/status/1394742784683298818}{https://twitter.com/Shayan86/status/1394742784683298818}.}
Thus, counting specific hashtag use over time allows to indirectly observe permanent suspensions on Twitter and there is a need for similar methodologies to be experimented on data from other platforms.
However, a point of concern is that data can also disappear because of other factors independent of platforms' content moderation.
To provide an example, McMinn et al. (2012)\cite{mcminn2013building} collected $120$ million Tweets related to general topics. 
The same published set of Tweet IDs was then used by Mazoyer et al. (2020)~\cite{mazoyer2020french}, to run the same collection in $2018$. Mazoyer et al. (2020)~\cite{mazoyer2020french} were only able to retrieve $67$ millions of Tweets (55\%). The authors argue that this drop is probably due to Tweets having been removed by users and Twitter accounts being shut by the users themselves.
As data can disappear from platforms over time, it is hard to distinguish whether platforms or users are the ones responsible for data removal. 
Hence in the special case of Twitter accounts related to QAnon, it is hard to identify how much of the drop in QAnon related hashtag use is due to the Twitter intervention or to accounts migrating to other platforms out of fear of having their accounts suspended by the platform or simply a drop in using the hashtag for other unidentified factors.

@@

A last concern regards the growing restrictions to access APIs (Perriam et al.(2019)~\cite{api}).
In April $2018$, likely as a response to the Cambridge Analytica scandal, Facebook severely limited access to their official APIs\footnote{See \href{https://about.fb.com/news/2018/04/restricting-data-access/}{https://about.fb.com/news/2018/04/restricting-data-access/} and \href{http://thepoliticsofsystems.net/2018/08/facebooks-app-review-and-how-independent-research-just-got-a-lot-harder/}{http://thepoliticsofsystems.net/2018/08/} \href{http://thepoliticsofsystems.net/2018/08/facebooks-app-review-and-how-independent-research-just-got-a-lot-harder/}{/facebooks-app-review-and-how-independent-research-just-got-a-lot-harder/}.} 
and CrowdTangle is today a reliable source for researchers to access Facebook data. 
However, it should be noted that CrowdTangle has been recently used by journalists to show that Facebook is disproportionally favoring highly partisan information\footnote{See the news article \href{https://www.economist.com/graphic-detail/2020/09/10/facebook-offers-a-distorted-view-of-american-news}{https://www.economist.com/graphic-detail/2020/09/10/facebook-offers-a-distorted-view-of-american-news} and the Tweets from \href{https://twitter.com/facebookstop10}{https://twitter.com/facebookstop10}. 
The last link is a Twitter account showing each day the ten most shared link on Facebook, most of them coming from extreme right and misinformation sources. 
Facebook has reacted to this here: \href{https://about.fb.com/news/2020/11/what-do-people-actually-see-on-facebook-in-the-us/}{https://about.fb.com/news/2020/11/what-do-people-actually-see-on-facebook-in-the-us/}.} 
and a general suspicion is growing that its access might be shut down in a near future to journalists and researchers.\footnote{See the news articles \href{https://www.nytimes.com/2021/07/14/technology/facebook-data.html}{https://www.nytimes.com/2021/07/14/technology/facebook-data.html} and \href{https://www.bloomberg.com/news/articles/2021-08-03/facebook-disables-accounts-tied-to-nyu-research-project}{https://www.bloomberg.com/news/articles/2021-08-03/facebook-disables-accounts-tied-to-nyu-research-project}.}
Regarding Twitter and YouTube, it is still possible and relatively easy to access their official APIs. %Nevertheless the number of Tweets that can be retrieved has decreased between the Twitter API v1 and v2. 
%In the v2, there is a monthly cap of $10,000,000$ Tweets for the academic research track\footnote{https://developer.twitter.com/en/docs/projects/overview},
%while in the v1 there was no monthly limit and only a limit on the number of requests that could be made in a 15-minute window\footnote{https://developer.twitter.com/en/docs/twitter-api/v1/rate-limits}.
%We thus noticed that we could get around $10$ times more Tweets per month with the v1 when compared to the v2 for one specific collection.

@@

In certain cases, the platform interventions were not motivated by content moderation to limit misinformation, but rather by other rule violations.
For example, the official reason for the suspension of Donald Trump's Twitter account was not misinformation\footnote{Although Donald Trump may have spread many misinformation during his governance, see for example: \href{http://allianceforscience.cornell.edu/blog/2020/10/what-drove-the-covid-misinformation-infodemic/}{http://allianceforscience.cornell.edu/blog/2020/10/what-drove-the-covid-misinformation-infodemic/}.}
but the violation of Twitter's Glorification of Violence Policy, with his last Tweets encouraging the Capitol's riot.\footnote{See \href{https://blog.twitter.com/en\_us/topics/company/2020/suspension}{https://blog.twitter.com/en\_us/topics/company/2020/suspension}.}
Similarly, when Facebook removed four pages related to the InfoWars website, the cited reason was repeated violations of Community Standards: 
``While much of the discussion around Infowars has been related to false news, which is a serious issue that we are working to address [...], none of the violations that spurred today’s removals were related to this''.\footnote{See the article \href{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}.}
Regarding The Beauty of Life website, its ban was officially due to coordinated inauthentic behavior,\footnote{See \href{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us/}{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us/}.} 
which includes using fake accounts that misrepresent one's identity or using methods to artificially boost the popularity of content. 
According to Facebook, coordinated inauthentic behavior is a distinct phenomenon from disinformation, as ``most of the content shared by coordinated manipulation campaigns isn’t probably false''.\footnote{See: \href{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}.} 
Nevertheless, both misinformation and coordinated inauthentic behavior were attested for the Beauty of Life, and reported to Facebook by the fact-checking organization Snopes.\footnote{See \href{ https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}, 
\href{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-} \href{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{fake-profiles/}  and 
\href{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}.} 
In the above examples, the compagnies' communication was very clear on the reason of the interventions (although one can suspect that the official reason given can depart from the real motivation behind an intervention). 
However, very often no official communication on behalf of the platforms can be found for a specific interventions observed in collected data. 
Because misinformation is often associated with creating fake accounts to speed its spread, and incitation to violence, it is sometimes unclear to distinguish between interventions due to misinformation and interventions due to other guideline violations, and more transparent communication on platforms' actions would help monitor the interventions specifically related to misinformation.

@@

{\bf A need for more consistent and adapted policies.} Misinformation on the web is a problem that is increasingly discussed since the $2016$ American elections and in the years that followed, platforms may have decided on a case-by-case basis which actions to enforce.
Infowars and The Beauty of Life domains have both several failed fact-checks\footnote{See \href{https://mediabiasfactcheck.com/infowars-alex-jones/}{https://mediabiasfactcheck.com/infowars-alex-jones/}  and \href{https://mediabiasfactcheck.com/the-bl/}{https://mediabiasfactcheck.com/the-bl/}.} and have both violated Facebook's community guidelines (incitation to violence and coordinated unauthentic behavior). 
Yet one's distribution was reduced for only half of a year, while the other one's links cannot be shared anymore on Facebook since the end of $2019$, and it is not clear why such different enforcements were applied to these comparable websites.
It is also possible that platforms usually follow specific rules on how to fight misinformation, but that sometimes the sudden media coverage of specific events leads them to take exceptional actions, such as with the Capitol riots and Trump's suspension from Facebook, Twitter and YouTube.
We would argue that policies should be clearly stated and carefully designed, rather than implementing interventions on a case-by-case basis and in response to the latest news events.

@@

Furthermore, it is likely that some interventions used to tackle online misinformation correspond to interventions that were designed for earlier policy areas.
To illustrate, YouTube explicitly states
\footnote{See \href{https://www.youtube.com/intl/en\_us/howyoutubeworks/our-commitments/fighting-misinformation/\#policies}{What policies exist to fight misinformation on Youtube?} (Last visited September 2021).}
that: ``Several policies in our Community Guidelines are directly applicable to misinformation'', such as the deceptive practices, impersonation and hate speech policies.
Hence interventions such as content deletion or account suspension, might be efficient for policy areas for which they were specifically designed, such as the policy area {\it Adult Nudity \& Sexual Activity}. Nevertheless, tackling online misinformation might need adapted interventions.
Namely, account suspensions or content deletion motivated by the spread of inaccurate information can be perceived as censorship by the targeted users. In particular, such interventions can generate platform migrations (see Rogers (2020)~\cite{rogers2020}) towards new alternative social media platforms (e.g. Telegram, Gab, Minds). 
\smallskip

Finally, the means to tackle online misinformation and understanding the phenomenon itself, is an active area of research. Developing pertinent interventions against misinformation depends on understanding the interventions currently used by platforms (e.g. for labeling, see Morrow et al. (2021)~\cite{morrow}) along with the psychology of interaction with social media (Pennycook and Rand (2021)~\cite{pennycook2021psychology}). 
Adding fact-checking labels to controversial content is certainly the intervention whose effectiveness was the most studied so far. 
It was shown that credibility indicators do reduce misperceptions and sharing, although their impact varied vastly depending on the presentation of the warnings (see Yaqub (2020)~\cite{yaqub2020effects} and Sharevski (2021)~\cite{sharevski2021misinformation}).
Moreover, a fact-checking label can lead to unpredicted consequences depending on the perception of the person reading it.
On Facebook, Twitter and YouTube, fact-checking messages are currently applied only to some content, while the rest of the content on a given platform remains naturally unlabelled. 
Such sparse application of warnings can lead to an ``implied truth'' effect, where users may assume that content without warnings have actually been verified and shown to be true (Pennycook et al. (2020)~\cite{pennycook2020implied}).
Using a different approach, Mosleh et al. (2021)~\cite{mosleh2021perverse} identified Twitter users sharing false news and replied to their false Tweets with links redirecting to fact-checking websites.
They observed a decrease in the quality and accuracy of the content subsequently shared by the corrected user. 
These research papers highlight the necessity to investigate potential unintended effects of well-meaning interventions. Furthermore, misinformation related interventions can be poorly misunderstood. Based on survey data, Saltz et al. (2021)~\cite{saltzbarrari} show that $40\%$ of the survey respondents (wrongly) believed that content is mostly or all fact-checked. To date, research about misinformation related interventions by platforms is still scarce and addressing the specific needs in terms of data for auditing and monitoring tasks can greatly help in improving our understanding of these interventions and study their impact.  
The methods presented in this paper can help to monitor not only the platforms' actions against misinformation, but also their subsequent consequences in real-life settings.
